---
title: "Correlation Analysis"
author: "Jenny Smith"
date: "April 14, 2017"
output: html_document
---

Purpose: To identify the genes that are covariates with MSLN expresssion. 

The Analysis Plan
1. subset the TPM expression data for the gene sets of interest (defined by Allison Kaeding)
2. Create correlation plots of these genes of interest (look how to highlight MSLN)
3. Perform spearman rank correlation of genes to MSLN (?) using cor.test and then multiple hypothesis testing to correct for # of genes. 

3. Perform linear modeling with Limma package to identify gene which positively and negatively correlated with MSLN and are significant. 
4. 

```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = 'H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/')
```


```{r message = FALSE, warning=FALSE}
library(stringr)
library(dplyr)
library(magrittr)
library(plyr)
library(ggplot2)
library(limma)
library(edgeR)
library(dendextend)
library(pvclust)
library(corrplot)
getwd()
```

```{r}
source("H:/scripts/ConversionScripts/Ensembl_ID_Convert_Function.r")
# source("H:/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/Heatmaps_Function.r")
```



#Define Functions 

##Subset Function
```{r}
subset_ExpnData <- function(matrix, genes){
  expnData <- t(matrix[which(rownames(matrix) %in% genes),]) #transpose and subset expression matrix
  sd <- apply(expnData,2, FUN = sd) #calc sd for all genes
  highSD <- sd[which(sd >= 1.0)] 
  expnData <- expnData[,names(highSD)] #select genes with sd > 1 to avoid zero sd in cor analysis
  return(expnData)
}
```



#Read in the data

#high Depth diagnostic samples
```{r}
HD_TPM <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/ExpressionData/TPM_withGeneSymbol_TARGET_AML.csv", stringsAsFactors = FALSE)
```

```{r}
Dx_HDTPM <- convert_Dx_Rx(HD_TPM, "ensemblSymbol", ID = "GENEID", subset = "Dx")

dim(Dx_HDTPM)
```

```{r}
Dx_HDTPM <- Dx_HDTPM[which(!duplicated(Dx_HDTPM$SYMBOL)), ]
Dx_HDTPM <- Dx_HDTPM[which(!is.na(Dx_HDTPM$SYMBOL)), ]
rownames(Dx_HDTPM) <- Dx_HDTPM$SYMBOL
Dx_HDTPM <- Dx_HDTPM[,-(1:2)]
```

```{r}
dim(Dx_HDTPM)
```


```{r}
Dx_HDTPM <- as.matrix(Dx_HDTPM)
head(Dx_HDTPM[,1:5])
# save(Dx_HDTPM, file = "TARGET_AML_HD_DxOnly_genesAsRowNames.RData")
```


```{r}
HD_counts <- get(load("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/ExpressionData/Dx_rawcounts_FilteredLowCounts_withGeneSymbol_TARGET_AML_Aug2014.RData"))
```

```{r}
HD_dge <- DGEList(HD_counts)
HD_dge <- 
HD_cpm <- 
```


#low Depth
```{r}
LD_TPM <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/ExpressionData/allTPM_BCCA28Apr2016.csv", stringsAsFactors = FALSE, row.names = 1)

LD_TPM <- LD_TPM[, which(! grepl("^BM", colnames(LD_TPM)))]
LD_TPM <- LD_TPM[which(! duplicated(LD_TPM$geneSymbol)), ]

rownames(LD_TPM) <- LD_TPM$geneSymbol
LD_TPM <- LD_TPM[,-1]
```

NOTE: Must determine why I have 466 samples in the CBFGLIS analysis and 476 here... 
```{r}
dim(LD_TPM)
```

```{r}
LD_TPM <- as.matrix(LD_TPM)
head(LD_TPM[,1:5])
# save(LD_TPM, file = "TARGET_AML_LD_DxOnly_genesAsRowNames.RData")
```

```{r}
LD_counts <- get(load("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/ExpressionData/combatAdjustedExp.RData"))

dim(LD_counts)
```




#Read in the genes (covariates) of  interest

#Top 400 highly expressed genes
```{r}
#genes selected by Allison Kaeding as beign highly expressed (top 400) 
highGenesdf <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/Correlation_analysis_w_MSLN.csv", stringsAsFactors = FALSE)

dim(highGenesdf)
```
```{r}
head(highGenesdf)
```

```{r}
highGenes <- highGenesdf$X
```

#CBF MLL associated genes
```{r}
CBF_MLL_Genedf <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/Gene_expression_associated_w_being_CBF_and_MLL.csv", stringsAsFactors = FALSE)

dim(CBF_MLL_Genesdf)
```

these are genes from literature searches. Allison Kaeding April 2017. 

```{r}
head(CBF_MLL_Genesdf)
```

```{r}
t.8.21 <- CBF_MLL_Genesdf$t.8.21.[-1]

inv.16 <- CBF_MLL_Genesdf$inv.16.[-1]

All.CBF <- CBF_MLL_Genesdf$All.CBF[-1]

KMT2A <- CBF_MLL_Genesdf$KMT2A[-1]
```

```{r}
genesToExamine <- list(highGenes, t.8.21, inv.16 , All.CBF, KMT2A)
names(genesToExamine) <- c("highGenes", "t.8.21", "inv.16" , "All.CBF", "KMT2A")
```

```{r}
summary(genesToExamine)
```

```{r}
names(genesToExamine)
```


```{r}
#For loop to add MSLN to each list
for (list in names(genesToExamine)){
  Newlist <- c(get(list), "MSLN")
  Newlist <- Newlist[order(Newlist)]
  # print(Newlist)
  genesToExamine[[list]] <- Newlist
}
```

```{r}
sapply(genesToExamine, length)
```


##Subset the TPM expression data for each set of genes
```{r}
HD_TPM_Expn <- lapply(genesToExamine, subset_ExpnData, matrix=Dx_HDTPM)

length(HD_TPM_Expn)
```

Not all the gene names are being found?? Must look into this.  
```{r}
sapply(HD_TPM_Expn, dim)
```

```{r}
LD_TPM_Expn <- lapply(genesToExamine, subset_ExpnData, matrix=LD_TPM)

length(LD_TPM_Expn)
```

```{r}
sapply(LD_TPM_Expn, dim)
```



#Define the correlation Matrices

Results: many many of these selected genes have very low variance and thus a SD near zero. 
This affects the correlation calculations for pearsons R. 

I edited the subset function to now have a sd of at least one or greater to avoid the error. 

Instead, I will try to use the WGCNA and Limma linear models to identify the most correlated genes.

```{r}
HD_corMat <- lapply(HD_TPM_Expn, FUN = cor)
LD_corMat <- lapply(LD_TPM_Expn, FUN = cor) 

```

```{r}
summary(HD_corMat)
```

```{r}
summary(LD_corMat)
```


#Create Correlation Plots
```{r}
# pdf(file="highGenes.pdf", height = 10, width = 10)
corrplot(HD_corMat$highGenes, method = "color", order = "FPC", tl.cex = 0.1, tl.col = "black", cl.pos="b", cl.cex = 0.75) #N = 160
# dev.off()
```


```{r fig.height=10, fig.width=10}
HD_corPlot <- lapply(HD_corMat, function(x) corrplot(x, method = "color", order = "FPC", tl.cex = 0.5, tl.col = "black", cl.pos="b", cl.cex = 0.75) )
```


```{r fig.height=10, fig.width=10}
LD_corPlot <- lapply(LD_corMat, function(x) corrplot(x, method = "color", order = "FPC", tl.cex = 0.5, tl.col = "black", cl.pos="b", cl.cex = 0.75))
```




