---
title: "Correlation Analysis"
author: "Jenny Smith"
date: "April 14, 2017"
output: html_document
---

Purpose: To identify the genes that are covariates with MSLN expresssion. 

The Analysis Plan
1. subset the TPM expression data for the gene sets of interest (defined by Allison Kaeding)
2. Create correlation plots of these genes of interest (look how to highlight MSLN)
3. Perform linear modeling with Limma package to identify gene which positively and negatively correlated with MSLN and are significant. 
4. WGCNA analysis

```{r setup}
library(knitr)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE, fig.align='center')
knitr::opts_knit$set(root.dir = 'H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/')
```


```{r message = FALSE, warning=FALSE}
library(stringr)
library(dplyr)
library(magrittr)
library(plyr)
library(ggplot2)
library(limma)
library(edgeR)
library(dendextend)
library(pvclust)
library(corrplot)
library(gtools)
getwd()
```

```{r}
source("H:/scripts/ConversionScripts/Ensembl_ID_Convert_Function.r")
# source("H:/scripts/RNAseq_Analysis/DifferentialExpn_PathwayAnalysis/Heatmaps_Function.r")
```



#Read in the data

#high Depth diagnostic samples
```{r}
HD_counts <- get(load("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/ExpressionData/Dx_rawcounts_FilteredLowCounts_withGeneSymbol_TARGET_AML_Aug2014.RData"))

dim(HD_counts)
```


#low Depth

```{r}
LD_counts <- get(load("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/ExpressionData/combatAdjustedExp.RData"))
LD_counts[which(LD_counts < 0, arr.ind=TRUE)] <- 0
LD_counts <- LD_counts[, which(! grepl("^BM", colnames(LD_counts)))]


dim(LD_counts)
```




#Read in the genes (covariates) of  interest

#Top 400 highly expressed genes
```{r}
#genes selectedas expressed only in AML by Rhonda 
# highGenesdf <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/Correlation_analysis_w_MSLN.csv", stringsAsFactors = FALSE)

highGenesdf <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/Genes_expressed_only_in_AML.csv", stringsAsFactors = FALSE)


dim(highGenesdf)
```

```{r}
head(highGenesdf)
```


```{r}
highGenes <- highGenesdf$X
```

#CBF MLL associated genes
```{r}
CBF_MLL_Genesdf <- read.csv("H:/RNA_seq_Analysis/2017.04.14_TARGET_AML_correlationMSLN/Gene_expression_associated_w_being_CBF_and_MLL.csv", stringsAsFactors = FALSE)

dim(CBF_MLL_Genesdf)
```

```{r}
head(CBF_MLL_Genesdf)
```

```{r}
t.8.21 <- CBF_MLL_Genesdf$t.8.21.[-1]

inv.16 <- CBF_MLL_Genesdf$inv.16.[-1]

All.CBF <- CBF_MLL_Genesdf$All.CBF[-1]

KMT2A <- CBF_MLL_Genesdf$KMT2A[-1]
```

```{r}
genesToExamine <- list(highGenes, t.8.21, inv.16 , All.CBF, KMT2A)
names(genesToExamine) <- c("highGenes", "t.8.21", "inv.16" , "All.CBF", "KMT2A")
```

```{r}
names(genesToExamine)
```

```{r}
sapply(genesToExamine, length)
```


#Define Functions to be used

```{r}
voom_Cor <- function(df, geneOfInterest, logCPM=NULL, voom=NULL){
  #df with count data, patient IDs are column names
  #ref: https://support.bioconductor.org/p/53554/

  require(edgeR)
  require(limma)

  dge <- DGEList(counts = df)
  keep.dge <- rowSums(cpm(dge) >= 1) >= (0.05*ncol(df)) #5% of dataset has cpm of at least 1 for a gene
  dge <- dge[keep.dge,] #subset for those genes with cmp >= 1 per gene
  
  if (is.null(logCPM)){
    dge <- calcNormFactors(dge) #calculate the normalization factors
    
  }else if (logCPM==TRUE){
    dge <- calcNormFactors(dge) #calculate the normalization factors
    log2Expn <- cpm(dge, log=TRUE, prior.count = 1)
    
  }else if (logCPM==FALSE){
    log2Expn <- apply(dge$counts, 2, function(x) log2(x + 1))
  }
 

  if (is.null(voom) && is.null(logCPM)){
    i <- which(rownames(dge$counts) == geneOfInterest)
    geneX <- dge$counts[i, ]
    Expnmat <- dge$counts[-i, ]
    
  } else if (exists("log2Expn") && is.null(voom)) {
    i <- which(rownames(log2Expn) == geneOfInterest)
    geneX <- log2Expn[i, ]
    Expnmat <- log2Expn[-i, ]
    
  } else if (voom==TRUE && is.null(logCPM)){
    v.lv <- voom(dge)
    i <- which(rownames(v.lv$E) == geneOfInterest)
    geneX <- v.lv$E[i, ]
    Expnmat <- v.lv[-i, ]
  }

  design <- cbind(1, geneX) #compare all genes to geneX expression (gene of interest)
  
  fit1 <- lmFit(Expnmat, design) #fit the linear model
  fit2 <- eBayes(fit1)
  
  table <- toptable(fit2, coef = 2, number = 5000, p.value = 0.05, adjust.method = "BH") #select top 5000 genes

  list <- list(dge, design, geneX, Expnmat, fit1, fit2, table)
  names(list) <- c("dge", "design", "GeneOfInterest_Expn","ExpnMatrix", "lmfit", "ebayesfit", "CorrGenes")

  return(list)
}
```



#Apply the linear modeling in the Voom method to MSLN
```{r}
HD_Vcor <- voom_Cor(HD_counts, "MSLN", voom = TRUE)

summary(HD_Vcor)
```

```{r}
dim(HD_Vcor$dge$counts)
```

```{r}
head(HD_Vcor$design)
```

```{r}
#selects the results from the contrast of geneX against all others (eg column 2 of design matrix results)
HD_Vcor$CorrGenes
```

```{r}
round(cor(HD_Vcor$GeneOfInterest_Expn, HD_Vcor$ExpnMatrix$E["TPPP3", ], method="pearson"), digits = 2) #not related to the linear regression 
```

```{r}
dim(HD_Vcor$ExpnMatrix$E)
```


```{r}
dim(HD_Vcor$ExpnMatrix$weights)
```


```{r}
# model <- lm(HD_Vcor$GeneOfInterest_Expn ~ HD_Vcor$ExpnMatrix$E["TPPP3", ]) #1.02280, R2 0.5799
model <- lm(HD_Vcor$ExpnMatrix$E["MED12L", ] ~ HD_Vcor$GeneOfInterest_Expn ) #0.56958, R2 0.5799

sum <- summary(model)

sum
```

```{r}
plot( HD_Vcor$ExpnMatrix$E["TPPP3", ], HD_Vcor$GeneOfInterest_Expn)
abline(model)
```

Regression coefficients represent the mean change in the response variable (MSLN) for one unit of change in the predictor variable (TPPP3) while holding other predictors in the model constant. 

```{r}
model$coefficients #this is the estimate for the TPPP3 expression from summary of the model fit
```

MSLN expression =  -0.9513849 + 1.0227994*TPPP3

for every unit (voom transformed read counts) increase in TPPP3 expression, you expect and an increase of 1.0227 read counts for MSLN expression. 

```{r}
HD_Vcor$fit$coefficients["TPPP3", ] #this is the fold-change from the linear regression= 1.480768 nonlog
```


The lmFit method uses lm.series to model fit (in this case expression of MSLN against all other genes, one at atime, for each individual). The lmFit method has a utility function lm.series that fits the expression matrix using ordinary least squares, with qr method.  

lm() base function fits the expression of MSLN to TPPP3 using ordinary least squares and qr method. 





```{r}
HD_Vcor$fit$p.value["TPPP3",] #p value is the pvalue for the linear model fitting between MSLN(geneX) and TPPP3. 
```


```{r}
LD_Vcor <- voom_Cor(LD_counts, "MSLN", voom=TRUE)

summary(LD_Vcor)
```

```{r}
dim(LD_Vcor$dge$counts)
```


```{r}
LD_Vcor$CorrGenes
```

```{r}
# write.csv(LD_Vcor$CorrGenes, file="correlatedGenes_MSLN_LimmaVoom.csv", row.names = TRUE)
```

```{r}
lm <- lm(LD_Vcor$GeneOfInterest_Expn ~ LD_Vcor$ExpnMatrix$E["TPPP3", ])

summary(lm)
```

```{r}
LD_Vcor$fit$coefficients["TPPP3", ] #foldchange is the coefficient
```

The coefficient for lmFit is 1.222 (0.2899085 in  log2 scale), while it is 1.266 in the ols method with lm(). 


```{r}
#Plot the voom  expression values 
plot(LD_Vcor$ExpnMatrix$E["TPPP3", ], LD_Vcor$GeneOfInterest_Expn, xlab="TPPP3 Expression", ylab="MSLN Expression")

abline(lm)
```

How do those outliers (far left clump), not affect the linear regression? 

```{r}
#residuals plots
plot(lm)
```


```{r}
cor(LD_Vcor$GeneOfInterest_Expn, LD_Vcor$ExpnMatrix$E["TPPP3", ], method = "kendall") #no relationship to correlation coefficients
round(LD_Vcor$CorrGenes["TPPP3", ]$logFC, digits = 4)

```


#Correlation using the limma trend method 

Use log2 TMMCPM normalized read counts and applied linear model in lmFit using ordinary least squares method to idenfify correlation bewteen each gene to MSLN. 

```{r}
HD_Log2Cor <- voom_Cor(HD_counts, "MSLN", logCPM = TRUE)

summary(HD_Log2Cor)
```


```{r}
dim(HD_Log2Cor$dge)
```

```{r}
HD_Log2Cor$CorrGenes
```



```{r}
#response = MSLN, predictor = TPPP3
model <- lm(HD_Log2Cor$GeneOfInterest_Expn ~ HD_Log2Cor$ExpnMatrix["TPPP3",])

plot(HD_Log2Cor$ExpnMatrix["TPPP3",], HD_Log2Cor$GeneOfInterest_Expn)
abline(model)
```

```{r}
summary(model) #adj. R2 is 0.5866 using log2CPM versus voom transfromation Adj. R-squared is 0.5799
```

```{r}
HD_Log2Cor$fit$coefficients["TPPP3", ] #1.489741 when converted out of log2
HD_Vcor$fit$coefficients["TPPP3", ] #1.480768
```



```{r}
LD_Log2Cor <- voom_Cor(LD_counts, "MSLN", logCPM = FALSE)

summary(LD_Log2Cor)
```

```{r}
LD_Log2Cor$CorrGenes
```

```{r}
LD_Log2Cor$fit$coefficients["TPPP3", ] # 1.24719
LD_Vcor$fit$coefficients["TPPP3", ] #1.222563
```

```{r}
lm <- lm(LD_Log2Cor$GeneOfInterest_Expn ~ LD_Log2Cor$ExpnMatrix["TPPP3", ])

plot(LD_Log2Cor$ExpnMatrix["TPPP3", ], LD_Log2Cor$GeneOfInterest_Expn)
abline(lm)
```



```{r}
summary(lm)
```


